A Vignetted View of Computer Science through Citation Analysis

### 2019-12-18 Scientometrics/comp.tex
lots of edits, look for your name, George

Noting that that the scientific literature serves a rich source of information to study the structure and historical development of a field, 
Salton and Bergmark conducted a historical study  in 1979 of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles)~\cite{salton_citation_1979}. These authors described the global structure of computer science as comprising three main areas (i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas noted were (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, and  (f) non-technical aspects, such as computer education.

Looking beyond Salton and Bergmark's historical triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS) published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field~\cite{acm_ref}. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications to the CCS, especially interdisciplinary articles or those from proximal fields, does not seem to be available.

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification} [add citations for Scopus and Web of Science]. Scopus and the Web of Science also include comprehensive bibliographies with citation links and proprietary unique identifiers for publications. 
All three systems rely on applying one or more journal-derived labels to articles.  
A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. 
Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. 
Others have also discussed and critiqued disciplinary assignments using journal-based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. 
Article classification systems have been constructed that escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, but do not enjoy widespread use. [Add reference to Scandinavian paper in QSS].

@@

 \subsection{Overview}
 In our study of the structure of a twenty year cohort of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the \emph{comp} dataset and mapped these clusters to journal-based categories.

 Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with direct citation being proposed as as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we used direct citation links as the basis for cluster formation, and also co-citation to obtain an alternative view.

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. The underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage may be incomplete and possibly biased. Cross-matching DBLP publications to records in the Scopus, abstract and citation database of peer-reviewed literature, enabled us to harvest the richer links in Scopus as well as extract links to publications from other disciplines. This cross-matching to Scopus, also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used DBLP articles from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications.

\subsection{Data}
A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Slightly over 95\% of the publications within were published after 1996. Publications were parsed from the XML source file and loaded into a PostgreSQL database. As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 87.8 million publications, into a custom schema in a PostgreSQL database. The total number of publications in Scopus labeled with major subject area of Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160. Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 nodes (publications and references) and 44,296,381 edges that represent citations within the dataset. This dataset is referred to as \emph{comp} (Fig~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}).

\subsection{Clustering methods} In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016}, of (i) the largest cluster should be no more than 10 times the smallest one, (ii) small clusters should be eliminated, (iii) small changes and replicates should yield similar results (``stability"),  (iv) computing time should be minimized where possible, and (v) the clustering should seem reasonable on a qualitative level (``intuitive sensibility").

@@

For an alternate view of these DBLP data we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles\cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973} and has been used for clustering publications. Co-citation, first described independently by Small and Marshakova in 1973, provides insight into the emergence of new ideas derived from the association of previously independent ones. Unlike clustering by direct citation with Graclus, where every input publication is assigned to a cluster and every citation is weighted equally, clustering by co-citation is progressively selective, involves assigning weights to co-citations, and also accounts for weak inter-cluster interactions~\cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}.

@@

In our implementation of variable level clustering (Fig.~\ref{fig:vlc_process}), we first calculated the number of citations accumulated across Scopus for all 2,685,356 articles in \emph{comp}. After discarding those publications without any citations, we restricted further analysis to highly cited articles--those in the 90th percentile of citations or higher--resulting in a dataset of size 212,311. We then identified 4,318,305 publications in Scopus that cite these 212,311 articles from \emph{comp}. For each of the 4,318,305 citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those in the set of 212,311 highly cited papers previously identified. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. A total of 46,463,117 unique co-cited pairs were thus obtained. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

We applied variable clustering with initial parameters of threshold ($t$) = median normalized co-citation frequency (quantile=0.5), increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster.  Using this approach, 22,232 clusters containing 84,591 nodes were generated, each containing less than 100 nodes. Clusters containing only 2 nodes were discarded bringing the total number of clusters down to 10,298. The publications in these 10,298 clusters were overwhelmingly drawn from the Physical Sciences (one of the 4 top level categories in the Scopus ASJC classification), of which computer science is a sub-category (Fig.~\ref{fig:quad-chart}).

@@

From the alternate perspective (columns), cluster 17 was the most diverse and contained publications annotated with 8 minor subject area labels: Biochemistry, Chemistry(all), Genetics, Molecular Biology,  Statistics \& Probability, Computational Theory \& Mathematics, and Computational Mathematics. Cluster 19 mapped to two areas but was excluded from qualitative analysis because of its high conductance value. Of the remaining clusters, the third (Cluster 2) represents interactions between the minor subject areas Theoretical Computer Science, Discrete Mathematics and Combinatorics, and Applied Mathematics as well as the more generic Computer Science (all). Clusters 3 and 4 include Computer Networks and Communications and Cluster 18 (Artificial Intelligence, Cognitive Neuroscience, and Neurology) and clusters 5--9 include Management Science, Operations Research, Information Systems, Modeling and Simulation, and Human-Computer Interaction.

These data suggest that areas central to computer science in 2019 (Salton and Bergmark's historical triad of hardware, software, and theory) are more likely to be found in multiple clusters than peripheral topics.
%, whereas areas such as biology, chemistry, and statistics are more restricted in distribution.  
A second inference is that, in some cases, journal-based classification and our article clusters align fairly well (Hardware and Architecture). 
A third inference is that the ASJC minor subject area ``Computer Applications" is relatively non-specific, and publications thus labeled are nearly ubiquitous (16/20).  Finally,   for the DBLP dataset, interactions with fields outside computer science (e.g., Biochemistry)  are  typically restricted to a single cluster or found in a small number of clusters.

We also note that these clusters cannot be adequately characterized by using a single CCS classification, a challenge that has been previously noted \cite{waltman_new_2012}.
For example, we
manually matched the top 25--50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, but not in all. For example, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization,  but 
the top cited papers often derived from biology and yet biology was clearly not  representative of all of these affected clusters.

\subsection{Clustering by Co-citation.} For an alternate examination of the data, we used co-citation frequencies to cluster the DBLP dataset as described above. Figure \ref{fig:heatmap_cocit} shows a heatmap in which clusters constructed by co-citation are mapped to Scopus ASJC minor subject areas. Mappings are only shown if they account for at least 15\% of the publications in a cluster (upper panel). Under these conditions, only 17 of 20 clusters mapped to minor subject areas where at least 15\% of the nodes in a cluster were labeled with a given subject area. When the threshold was reduced to 10\% all 20 clusters indicated at least one mapping to a minor subject area. We interpret these data as indicative of two kinds of co-citational linkages (i) intra-cluster (ii) cross cluster. The latter predominates and is detected at lower thresholds (lower panel). No cluster maps to more than 3 minor subject area at the 15\% threshold in comparison to a maximum of 8 minor subject areas for clustering by direct citation (Fig 3, Cluster 17). It should be noted, however, that the size of clusters constructed by co-citation is much smaller than those generated by direct citation through Graclus.

@@

\section{Discussion} In this article, we have tried to revisit a characterization of computer science from 1979~\cite{salton_citation_1979} in the light of expansion and diversification of this field. We analyzed considerably more data, 2.68 million publications versus 391, and leveraged access to two bibliographic databases, DBLP and Scopus. The former identifies a sample of the scientific literature that is computer science and the latter (roughly 10 times larger at 87.XX million publications) enabled analyzing citations across the four major groups of Physical Science, Social Sciences, Life Sciences, and Health Sciences in the Scopus ASJC classification.

Obvious critiques of our approach are (i) use of the DBLP dataset, which may not capture all aspects of computer science and its interactions with other fields (ii) partitioning DBLP into a weakly justified number of clusters and arbitrarily defined thresholds for detection (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. Our focus on high-level features may mitigate these valid concerns and preference for simplicity and intuitiveness in our choice of methods.

Reconciling direct citation clusters to the Scopus ASJC classification, yielded results that are consistent with the observation of Waltman and van Eck (2012)~\cite{waltman_new_2012}, `that traditional disciplines such as those just mentioned only partly reflect the actual organization of today’s scientific research'. Of interest to us, was the single obviously multidisciplinary cluster in which at least 15\% of its component publications were labeled with the ASJC minor subject areas Biochemistry, Chemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, Genetics, Molecular Biology, Statistics \& Probability followed by the cluster mapping to Artificial Intelligence, Cognitive Neuroscience, and Neurology.  Perhaps at the opposite end of this spectrum is the more insular cluster that maps to Hardware and Architecture, Electrical and Electronic Engineering, and Software. To us, these data suggest a diverse field of activities [not sure what that means].

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.4]{graclus_comparison.pdf}
% figure caption is below the figure
\caption{Conductance Measurements of Clusters Generated by Graclus.  2,685,356 DBLP publications, 7,129,006 cited references, and 44,296,381 citations were clustered using Graclus into 18 (grac\_18), 20 (grac\_20), or 22 (grac-22) clusters. Conductance, $\phi(S)$, was measured for these clusters considering only the edges between publications using the formula: $\phi(S)=|\partial(S)|/min(vol(S),2m-vol(S)$, where $\partial(S)$ is the boundary (number of edges leaving a set), vol(S) is volume of a set of vertices as the sum of the degrees of the vertices in a set, and $m$ is the number of undirected edges in a set~\cite{shun_parallel_2016}.}
\label{fig:graclus_comparison_conductance}       
\end{figure}

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{scopus_dblp_heatmap3.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by direct citation. The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least 15\% of the publications in the cluster.
 }
\label{fig:heatmap_cocit}       % Give a unique label
\end{figure}
\newpage

>>>>

Noting that that the scientific literature serves a rich source of information to study the structure and historical development of a field, 
Salton and Bergmark conducted a historical study  in 1979 of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles)~\cite{salton_citation_1979}. These authors described the global structure of computer science as comprising three main areas: (i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas noted were (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, and  (f) non-technical aspects, such as computer education.

Looking beyond Salton and Bergmark's historical triad of theoretical foundations, hardware and computer systems, and software, the Computing Classification System (CCS) published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field~\cite{acm_ref}. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications to the CCS, especially interdisciplinary articles or those from proximal fields, does not seem to be available.

Other classification systems are available, such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification} [add citations for Scopus and Web of Science]. Scopus and the Web of Science also include comprehensive bibliographies with citation links and proprietary unique identifiers for publications. 
All three systems rely on applying one or more journal-derived labels to articles.  
A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. 
Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) \cite{CSCD} and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. 
Others have also discussed and critiqued disciplinary assignments using journal-based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. 
Article classification systems have been constructed that escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, but do not enjoy widespread use. [Add reference to Scandinavian paper in QSS].

@@

 \subsection{Overview}
 In our study of the structure of a twenty year cohort of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the  computer science literature (as represented in our \emph{comp} dataset, described below) and mapped these clusters to journal-based categories.

 Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation, with direct citation being proposed as as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we used direct citation links as the basis for cluster formation, and also co-citation to obtain an alternative view.

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. The underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage may be incomplete and possibly biased. Cross-matching DBLP publications to records in the Scopus abstract and citation database of peer-reviewed literature, enabled us to harvest the richer links in Scopus as well as extract links to publications from other disciplines. This cross-matching to Scopus also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used DBLP articles from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications.

\subsection{Data}
A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Slightly over 95\% of the publications within were published after 1996. Publications were parsed from the XML source file and loaded into a PostgreSQL database. As part of implementing a larger data platform for research evaluation~\cite{GithubERNIE2019}, we have previously parsed the Scopus dataset, presently at over 87.8 million publications, into a custom schema in a PostgreSQL database. The total number of publications in Scopus labeled with major subject area of Computer Science (in turn a subset of the Physical Sciences subject area) is 5,835,160. Records in the DBLP dataset were matched to Scopus identifiers using digital object identifiers (DOIs). This procedure resulted in a dataset of 2,685,356 DBLP publications with Scopus identifiers where 1,278,322 (47.6\%) were labeled as article and 1,407,034 (52.4\%) as conference proceedings.  References cited by these publications were then extracted from Scopus (7,129,006 records), resulting in a total of 8,000,411 
publications and references. 
We represented these data using a graph where the 8,000,411 nodes represent
publications and references  and the 44,296,381 undirected edges represent citations within the dataset. 
We refer to these data as the  \emph{comp} dataset (Fig~\ref{fig:ar_cp_annotation}, Table~\ref{tab:comp}).

\subsection{Clustering methods} In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016} that (i) the largest cluster should be no more than 10 times the smallest one, (ii) small clusters should be eliminated, (iii) small changes and replicates should yield similar results (``stability"),  (iv) computing time should be minimized where possible, and (v) the clustering should seem reasonable on a qualitative level (``intuitive sensibility").

@@

For an alternate view of these DBLP data, we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles \cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973}. % and has been used for clustering publications. %George, add citation 
Co-citation, first described independently by Small and Marshakova in 1973 \cite{marshakova-shaikevich_co-citation_1973}, provides insight into the emergence of new ideas derived from the association of previously independent ones. Unlike clustering by direct citation, where every input publication is assigned to a cluster and every citation is weighted equally, the co-citation relationship between papers is weighted to represent the strength of the co-citation history.
Because this produces a weighted graph, clustering methods that address weights are required.
Furthermore, co-citation produces relatively weak inter-cluster interactions, requiring additional modifications to standard clustering approaches \cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}.

@@

In our implementation of variable level clustering (Fig.~\ref{fig:quad-chart}(a)), we first calculated the number of citations accumulated across Scopus for all 2,685,356 articles in \emph{comp}. After discarding those publications without any citations, we restricted further analysis to highly cited articles--those in the 90th percentile of citations or higher--resulting in a dataset of size 212,311. We then identified 4,318,305 publications in Scopus that cite these 212,311 articles from \emph{comp}. For each of the 4,318,305 citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. 
The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

We applied variable clustering with initial parameters as follows: the threshold defining the initial median normalized co-citation frequency (quantile) $t=0.5$, increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster.  Using this approach, 22,232 clusters containing 84,591 nodes were generated, each containing less than 100 nodes. Clusters containing only 2 nodes were discarded, bringing the total number of clusters down to 10,298. The publications in these 10,298 clusters were overwhelmingly drawn from the Physical Sciences (one of the 4 top level categories in the Scopus ASJC classification), of which computer science is a sub-category (Fig.~\ref{fig:quad-chart}).

@@

From the alternate perspective (columns), Cluster 17 was the most diverse and contained publications annotated with 8 minor subject area labels: Biochemistry, Chemistry(all), Genetics, Molecular Biology,  Statistics \& Probability, Computational Theory \& Mathematics, and Computational Mathematics. Cluster 19 mapped to two areas but was excluded from qualitative analysis because of its high conductance value. Of the remaining clusters, Cluster 2 represents interactions between the minor subject areas Theoretical Computer Science, Discrete Mathematics and Combinatorics, and Applied Mathematics as well as the more generic Computer Science (all). Clusters 3 and 4 include Computer Networks and Communications, and Cluster 18 (Artificial Intelligence, Cognitive Neuroscience, and Neurology) and clusters 5--9 include Management Science, Operations Research, Information Systems, Modeling and Simulation, and Human-Computer Interaction.

These data suggest that areas central to computer science in 2019 (Salton and Bergmark's historical triad of hardware, software, and theory) are more likely to be found in multiple clusters than peripheral topics.
%, whereas areas such as biology, chemistry, and statistics are more restricted in distribution.  
A second inference is that, in some cases, journal-based classification and our article clusters align fairly well (Hardware and Architecture). 
A third inference is that the ASJC minor subject area ``Computer Applications" is relatively non-specific, and publications thus labeled are nearly ubiquitous (present in 16 out of 20 clusters).  Finally,   for the DBLP dataset, interactions with fields outside computer science (e.g., Biochemistry)  are  typically restricted to a single cluster or found in a small number of clusters.

We also note that these clusters cannot be adequately characterized by using a single CCS classification, a challenge that has been previously noted \cite{waltman_new_2012}.
For example, we
manually matched the top 25--50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, but not in all. For example, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization,  but 
the top cited papers often derived from biology and yet biology was clearly not  representative of any of these affected clusters.

\subsection{Clustering by Co-citation.} 
For an alternate examination of the data, we used co-citation frequencies to cluster the DBLP dataset as described above. Figure \ref{fig:heatmap_cocit} shows a heatmap in which clusters constructed by co-citation are mapped to Scopus ASJC minor subject areas, with the top subfigure showing results for those labels that
account for at least 15\% of the publications in a cluster, and the bottom subfigure showing results for those labels that account for at least 10\% of the publications.
%Mappings are only shown if they account for at least 15\% of the publications in a cluster (upper panel). 
Under these conditions, only 17 of 20 clusters mapped to minor subject areas where at least 15\% of the nodes in a cluster were labeled with a given subject area. When the threshold was reduced to 10\%, all 20 clusters indicated at least one mapping to a minor subject area. We interpret these data as indicative of two kinds of co-citational linkages (i) intra-cluster (ii) cross cluster. The latter predominates and is detected at lower thresholds (lower panel). No cluster maps to more than three minor subject area at the 15\% threshold,  in contrast to a maximum of eight minor subject areas for clustering by direct citation (Fig. 3, Cluster 17). It should be noted, however, that the size of clusters constructed by co-citation is much smaller than those generated by direct citation through Graclus. 
%George, add some comments and discussion

@@

\section{Discussion} In this article, we revisited a characterization of computer science from 1979~\cite{salton_citation_1979} in the light of expansion and diversification of this field. We analyzed considerably more data, 2.68 million publications versus 391, and leveraged access to two bibliographic databases, DBLP and Scopus. The former identifies a sample of the scientific literature that is computer science and the latter (roughly 10 times larger at 87.XX million publications) enabled analyzing citations across the four major groups of Physical Science, Social Sciences, Life Sciences, and Health Sciences in the Scopus ASJC classification.

Obvious critiques of our approach are: (i) use of the DBLP dataset, which may not capture all aspects of computer science and its interactions with other fields, (ii) partitioning DBLP into a weakly justified number of clusters and arbitrarily defined thresholds for detection, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. 
Our focus on high-level features and preference for simplicity and intuitiveness in our choice of methods may mitigate these concerns.

Reconciling direct citation clusters to the Scopus ASJC classification yielded results that are consistent with the observation of Waltman and van Eck (2012)~\cite{waltman_new_2012}, `that traditional disciplines such as those just mentioned only partly reflect the actual organization of today’s scientific research'. Of interest to us, was the single obviously multidisciplinary cluster in which at least 15\% of its component publications were labeled with the ASJC minor subject areas Biochemistry, Chemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, Genetics, Molecular Biology, Statistics \& Probability, followed by the cluster mapping to Artificial Intelligence, Cognitive Neuroscience, and Neurology.  Perhaps at the opposite end of this spectrum is the more insular cluster that maps to Hardware and Architecture, Electrical and Electronic Engineering, and Software. To us, these data suggest that some subfields within computer science are primarily inward looking (remain concerned with fundamental questions in computer science), while others are more actively engaged with fields external to computer science.

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.4]{graclus_comparison.pdf}
% figure caption is below the figure
\caption{Conductance Measurements of Clusters Generated by Graclus of the direct citation dataset.  2,685,356 DBLP publications, 7,129,006 cited references, and 44,296,381 citations were clustered using Graclus into 18 (grac\_18), 20 (grac\_20), or 22 (grac-22) clusters. Conductance, $\phi(S)$, was measured for these clusters considering only the edges between publications using the formula: $\phi(S)=|\partial(S)|/min(vol(S),2m-vol(S)$, where $\partial(S)$ is the boundary (number of edges leaving a set), vol(S) is volume of a set of vertices as the sum of the degrees of the vertices in a set, and $m$ is the number of undirected edges in a set~\cite{shun_parallel_2016}.}
\label{fig:graclus_comparison_conductance}       
\end{figure}

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.45]{scopus_dblp_heatmap3.pdf}
% figure caption is below the figure
\caption{Heat map for the clustering obtained by co-citation, using two thresholds for inclusion (top: 15\%, bottom: 10\%). The  y-axis (rows) correspond to topics, defined by Scopus characterizations, and the
x-axis (columns) represent the 20 different clusters.  
Each cluster is characterized by topics that label at least the required minimum percentage of publications in the cluster (top: 15\%, bottom: 10\%).   }
\label{fig:heatmap_cocit}       % Give a unique label
\end{figure}
\newpage
